---
title: "Introduction to timbr"
knitr:
  opts_chunk:
    collapse: true
    comment: '#>'
---

```{r}
#| label: setup
#| message: false
library(timbr)
library(dplyr)
```

Hierarchical data usually lives in a flat data frame with one column per
level: sales by region, store, and product; energy demand by region, sector,
and fuel; cost breakdowns by assembly and part. dplyr handles one level of
aggregation naturally, but each `summarise()` returns a new, flat data frame.
After aggregating, you no longer know how the levels relate to each other, and
computations that need both parents and children -- rollups with losses,
top-down allocations, scenario propagation -- turn into a dance of repeated
joins.

timbr keeps the whole hierarchy in a single object called a **forest**. A
forest is built from a data frame, prints like one, and supports familiar
dplyr verbs -- but its rows are nodes of trees, so aggregating adds parent
nodes instead of throwing the structure away.

This article walks through the basics with a small energy demand dataset:

```{r}
energy <- tribble(
  ~region , ~sector     , ~fuel         , ~demand ,
  "north" , "household" , "electricity" ,      20 ,
  "north" , "household" , "gas"         ,      25 ,
  "north" , "industry"  , "electricity" ,      30 ,
  "north" , "industry"  , "gas"         ,      45 ,
  "south" , "household" , "electricity" ,      40 ,
  "south" , "household" , "gas"         ,      15 ,
  "south" , "industry"  , "electricity" ,      60 ,
  "south" , "industry"  , "gas"         ,      30
)
```

## From data frame to forest

`forest_by()` works like `dplyr::group_by()`: you list the columns that define
the hierarchy, from the top level down. The last column becomes the nodes of
the forest, and the remaining columns become groups:

```{r}
fr <- energy |>
  forest_by(region, sector, fuel)
fr
```

The `.` column shows each node's name and value (here, `fuel` nodes). The
other columns -- called **features** -- hold the data attached to each node,
like `demand` here.

## Hierarchical aggregation

`summarise()` on a forest looks just like `summarise()` on a grouped data
frame. The difference is that instead of discarding the `fuel` level, it adds
parent nodes on top of it:

```{r}
fr_sector <- fr |>
  summarise(demand = sum(demand))
fr_sector
```

The header now shows the tree structure: `sector` nodes with `fuel` children.
Summarise again to add a `region` level:

```{r}
fr_region <- fr_sector |>
  summarise(demand = sum(demand))
fr_region
```

The forest prints its root nodes, but every level is still there -- nothing
has been lost along the way.

## Moving around the hierarchy

`children()` steps down one level, turning the current roots into groups:

```{r}
children(fr_region)
```

`climb()` jumps to a named level anywhere in the forest:

```{r}
fr_region |>
  climb(fuel)
```

And `leaves()` returns the terminal nodes:

```{r}
leaves(fr_region)
```

## Combining forests

Real hierarchies are often ragged: different branches have different levels.
Suppose transport demand is broken down by mode instead of sector and fuel.
`rbind()` combines forests, and a further `summarise()` joins them under a
common root:

```{r}
transport <- tribble(
  ~region , ~mode , ~demand ,
  "north" , "car" ,      35 ,
  "north" , "bus" ,      10 ,
  "south" , "car" ,      45 ,
  "south" , "bus" ,      15
)

fr_transport <- transport |>
  forest_by(region, mode)

fr_total <- rbind(fr_sector, fr_transport) |>
  summarise(demand = sum(demand))
fr_total
```

Each `region` node now has two differently shaped subtrees, and hierarchical
operations keep working across both.

## Custom hierarchical computations with traverse()

`summarise()` covers computations where each parent is a simple aggregate of
its children. `traverse()` handles everything else: it applies a function at
every parent-children pair, walking the forest from the leaves up. For
example, suppose demand at each aggregation level includes 5% distribution
losses on top of its children's total:

```{r}
fr_total |>
  traverse(function(parent, children) {
    parent$demand <- sum(children$demand) * 1.05
    parent
  })
```

Because `traverse()` walks the tree level by level, the losses compound
naturally: regional totals include losses on losses. With `.climb = TRUE` the
walk runs in the opposite direction, from roots to leaves, and the function
receives `(children, parent)` -- useful for top-down allocations such as
distributing a target or an overhead cost across children.

## Back to data frames

`as_tibble()` converts the root nodes of a forest back into an ordinary
tibble:

```{r}
fr_sector |>
  as_tibble()
```

Combined with `climb()` or `leaves()`, this lets you pull out any level of
the hierarchy once the computation is done.

## Learn more

- The [function reference](https://uchidamizuki.github.io/timbr/reference/)
  covers the remaining verbs, including `mutate()`, `select()`, `relocate()`,
  `rows_update()`, and `rows_patch()` for forests.
- Forests convert to tidygraph objects via `as_tbl_graph()` for
  graph-oriented workflows.
